Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-05-24 paper-amia-2018/manuscript.tex
update result section

As follows from Table~\ref{tab:result_base}, KNN outperforms all other models in terms of precision and F1-measure achieving 0.808 precision with 0.728 F1-measure for new segment detection. KNN also shows superior performance in all performance metrics for ``same segment'' class and wighted average over ``new segment'' and ``same segment'' classes. NB demonstrates the lowest performance among all models in terms of all performance metrics. On the other hand, SVM has the highest recall of 0.673 when only lexical features are used to identify ``new segment''. Results indicate that performance of classifiers is remarkably higher for ``same segment'' class compared to ``new segment'' class, which is expected, since 96.71\% instances belong to ``same segment'' categories. For example, KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, in same segment identification compared to new segment detection. When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for new segment identification, 0.1\% higher precision than LSTM for same segment identification and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification. \\

@@

Table~\ref{tab:result_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``new segment'' class. We observed that performance is remarkably better for ``new segment'' class compared to new segment detection, which is expected, since 96.71\% instances belong to ``same segment'' class. GRU demonstrates the same result as KNN for precision and F1-measure, but not for recall. SVM shows moderate performance, achieving 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively, when lexical features are used. \\

@@

As shown in Table~\ref{tab:result_not_boundary}, performance of classifiers is remarkably higher for ``same segment'' class compared to new segment class, which is expected, since 96.71\% instances belong to ``same segment'' categories. KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, for lexical features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure, respectively, for combined features compared to new segment classification. In contrast to new segment detection, RNN demonstrates the highest performance among all models. The impact of punctuation and topic model-based features is consistent with ``new segment'' classification. Results indicate that F1-measure increases by 0\%, and 0.1\% for SVM and KNN models and decreases by 0.1\% for NB. \\

>>>>

As follows from Table~\ref{tab:result_base}, KNN outperforms all other models in terms of precision and F1-measure achieving 0.808 precision with 0.728 F1-measure for new segment detection. KNN also shows superior performance in all performance metrics for ``same segment'' class and wighted average over ``new segment'' and ``same segment'' classes. NB demonstrates the lowest performance among all models in terms of all performance metrics. On the other hand, SVM has the highest recall of 0.673 when only lexical features are used to identify ``new segment''. Results indicate that performance of classifiers is remarkably higher for ``same segment'' class compared to ``new segment'' class, which is expected, since 96.71\% instances belong to ``same segment'' categories. For example, KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, in same segment identification compared to new segment detection. When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for new segment identification, 0.1\% higher precision than LSTM for same segment identification, and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification. \\

@@

Table~\ref{tab:result_boundary} summarizes the performance of NB, SVM and KNN models for detecting ``new segment'' class. We observed that classification performance significantly improved for ``new segment'' class when lexical features are used in combination with punctuation and model-based topic features. Precision increases by 4.99\% and 0\%, 11.29\% and 0.74\%, and 9.71\% and 0.74\% for SVM and KNN; recall increases by 18.72\% and 17.06\%, 26.15\% and 15.64\%, and 24.67\% and 16.90\% for SVM and KNN; and F1-measure increases by 11.75\% and 6.18\%, 18.60\% and 5.91\%, and 17.06\% and 6.46\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively. On the other hand, NB has very negligable impact of punctuation and model-based topic features. In contrast to results in Table~\ref{tab:result_base}, SVM demonstrates the best performance among all models achieving 0.848, 0.849, and 0.848 for precision, recall, and F1-measure, respectively, when DL-LDA-based topic fetaures are used in conjunction with lexical and punctuation features. NB and KNN provide their best performance when LCA-based topic features are used, achieving 0.626 and 0.775 F1-measure, respectively.\\

@@

Similar to baseline performance, classification performance is remarkably higher for same segment class compared to new segment class (Table~\ref{tab:result_not_boundary}). The impact of punctuation and topic model-based features is also consistent with ``new segment'' classification while improvements are not as significant as ``new segment'' class. We observed that precision increases by 4.99\% and 0\%, 11.29\% and 0.74\%, and 9.71\% and 0.74\% for SVM and KNN; recall increases by 18.72\% and 17.06\%, 26.15\% and 15.64\%, and 24.67\% and 16.90\% for SVM and KNN; and F1-measure increases by 11.75\% and 6.18\%, 18.60\% and 5.91\%, and 17.06\% and 6.46\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively. Similar as Table \ref{tab:result_boundary}, NB and KNN demonstrate the lowest and moderate performance on the task of identifying same segment, respectively. \\
